---
title: "Lab 03 - Nobel laureates"
author: "Lilly McClendon"
date: "01-30-2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r skim-data}
library(skimr)
skim(nobel)
print(nobel)
```

1. There are 935 observations, and there are 25 variables in the dataset (26 variables if you include the ID column). Each row represents a different nobel prize winning individual. 


### Exercise 2
```{r nobel-living-data-frame-filter, eval=TRUE}
library(dplyr)
nobel_living <- nobel %>% 
  filter(!is.na(country)) %>% 
  filter(gender != "org") %>% 
  filter(is.na(died_date))
nrow(nobel_living)
```


### Exercise 3
```{r country-us-mutate}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r limit-prize-categories}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r faceted-bar-plot}
nobel_living_science %>% 
  ggplot(aes(x=country_us)) + 
  geom_bar() + 
  ggtitle("Relationship between the Category of Prize and Country when Prize Won") + 
  facet_wrap(~category) + 
  coord_flip()
```

#### Visualization Interpretation 

For the Nobel Prize categories of chemistry, economics, medicine, and physics, more prizes have been awarded when the individual resides in the U.S. However, there is a greater difference for some categories than others. There seems to be quite a large proportion of economic Nobel prizes awarded to individual's currently residing in the U.S. Given this faceted bar plot, I would say that the Buzzfeed data seems to be supported by the data. 


### Exercise 4 

```{r born_country_us-mutate, eval=TRUE}
 nobel_living_science_born <- nobel_living_science %>% 
  mutate( 
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  ) 
```

```{r filter_bornUS-resideUS_test, eval=TRUE}
nobel_living_science_born %>% 
  filter(
    born_country_us == "USA", 
    country_us == "USA", 
  ) %>%
  select(born_country_us, country_us)
```

#### Number of winners born in the US and won while residing in the US

Of the US-based nobel laureates, 102 were also born in the US. 


### Exercise 5

```{r plot-bornus-resideus, eval = TRUE}
ggplot(nobel_living_science_born, aes(x=country_us, fill=born_country_us, color=born_country_us)) + 
  geom_bar(stat = "count") +
  ggtitle("Relationship between the Category of Prize, Country when Prize Won, and Born Country") +
  facet_wrap(~category) + 
  coord_flip()
```

#### Visualization of Country Born and Country Reside 

After adding a second variable for country of birth to the visualization, the data appears to support Buzzfeed's claim that immigration is important for science in the US, however, there are a fair number of US based laureates who were born in the US as well. I would say that the data supports Buzzfeed's claim but it is not as drastic as they are saying. 


### Exercise 6
```{r filter-count-resideUS-bornOther, eval=TRUE}
nobel_living_science_born %>%
  filter(country == "USA" & born_country != "USA") %>%
  count(born_country, sort = TRUE)
```

#### Most common born country but won in US

Germany and the United Kingdom both had 7 people born outside of the US, but won their nobel prize while residing in the US. 
