---
title: "Lab 03 - Nobel laureates"
author: "Lilly McClendon"
date: "01-30-2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r skim-data}
library(skimr)
skim(nobel)
print(nobel)
```

1. There are 935 observations, and there are 25 variables in the dataset (26 variables if you include the ID column). Each row represents a different nobel prize winning individual. 


```{r nobel-living-data-frame-filter, eval=TRUE}
library(dplyr)
nobel_living <- nobel %>% 
  filter(!is.na(country)) %>% 
  filter(gender != "org") %>% 
  filter(is.na(died_date))
nrow(nobel_living)
```


### Exercise 2
```{r country-us-mutate}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r limit-prize-categories}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r faceted-bar-plot}
nobel_living_science %>% 
  ggplot(aes(x=country_us)) + 
  geom_bar() + 
  ggtitle("Relationship between the Category of Prize and Country when Prize Won") + 
  facet_wrap(~category) + 
  coord_flip()
```

#### Visualization Interpretation 

For the Nobel Prize categories of chemistry, economics, medicine, and physics, more prizes have been awarded when the individual resides in the U.S. However, there is a greater difference for some categories than others. There seems to be quite a large proportion of economic Nobel prizes awarded to individual's currently residing in the U.S. Given this faceted bar plot, I would say that the Buzzfeed data seems to be supported by the data. 

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
