---
title: "Lab 03 - Nobel laureates"
author: "Lilly McClendon"
date: "01-30-2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r skim-data}
library(skimr)
skim(nobel)
print(nobel)
```

1. There are 935 observations, and there are 25 variables in the dataset (26 variables if you include the ID column). Each row represents a different nobel prize winning individual. 


```{r nobel-living-data-frame-filter, eval=TRUE}
library(dplyr)
nobel_living <- nobel %>% 
  filter(!is.na(country)) %>% 
  filter(gender != "org") %>% 
  filter(is.na(died_date))
nrow(nobel_living)
```


### Exercise 2

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
